VIPOS: Virtualized Intermittently Powered Runtime Environment (``VIPRE''?)

### 2017-06-29 paper/paper.tex
rough draft of the intro

\usepackage{graphicx}  
\usepackage{subfigure}
\newcommand{\TODO}[1]{\bf \em \textcolor{red}{TODO: #1}\xspace}

@@

\title{VIPOS: Virtualized Intermittently Powered Runtime Environment (``VIPRE''?)}

% \author{
% 	\IEEEauthorblockN{Amjad Yousef Majid, Przemys{\l}aw Pawe{\l}czak}
% 	\IEEEauthorblockA{TU Delft, Mekelweg\,4, 2628\,CD Delft, NL\\
% 	\{a.y.majid, p.pawelczak\}@tudelft.nl}
% }

@@

\section{abstract}
Enabling battery-free devices is a mandatory step towards the environment friendly IoT. However, removing these large energy reservoirs (the batteries) makes the devices operate on an intermittent power supply, which make sustaining any long computation very challenging. correspondingly, there are two proposed approaches to enable long-running computations on intermittenly powered devices (IPDs): (i) Checkpointing based approach, where the volatile state of a program is frequently saved to the non-volatile memory and; (ii)  Task based approach, where a programmer splits the code into small idempotent modules. The Task based approach shows a better performance than the checkpointing approach. However, the task based method requires that the energy to execute any task must not exceed the maximum limit of the energy buffer. This limitation makes this solution a customized one that suits a particular energy buffer size.

@@

\section{Introduction}
\label{sec:intro}

@@

\TODO{Don't forget all the references in this section!}
New programming and execution models~\cite{dino,ratchet,chain,alpaca} and hardware
mechanisms~\cite{clank, idetic} help avoid data inconsistency

>>>>

\usepackage{graphicx}  
\usepackage{subfigure}
\newcommand{\TODO}[1]{{\bf \em \textcolor{red}{TODO: #1}\xspace}}
\newcommand{\sys}{NAME\xspace}

@@

\title{VIPOS: Virtualized Intermittently Powered Runtime Environment (``VIPRE''?)}
% \author{
% 	\IEEEauthorblockN{Amjad Yousef Majid, Przemys{\l}aw Pawe{\l}czak}
% 	\IEEEauthorblockA{TU Delft, Mekelweg\,4, 2628\,CD Delft, NL\\
% 	\{a.y.majid, p.pawelczak\}@tudelft.nl}
% }

@@

\section{abstract}
\TODO{Need to re-write this to make it consistent with the story after we have the story in place}
Enabling battery-free devices is a mandatory step towards the environment friendly IoT. However, removing these large energy reservoirs (the batteries) makes the devices operate on an intermittent power supply, which make sustaining any long computation very challenging. correspondingly, there are two proposed approaches to enable long-running computations on intermittenly powered devices (IPDs): (i) Checkpointing based approach, where the volatile state of a program is frequently saved to the non-volatile memory and; (ii)  Task based approach, where a programmer splits the code into small idempotent modules. The Task based approach shows a better performance than the checkpointing approach. However, the task based method requires that the energy to execute any task must not exceed the maximum limit of the energy buffer. This limitation makes this solution a customized one that suits a particular energy buffer size.

@@

\section{Introduction}
\label{sec:intro}
\TODO{VIPER: Virtualized Intermittent Programming model, Execution model, and Runtime environment}
\TODO{PREVIEW: Programming and Runtime Environment with Virtualized Intermittent Execution Windows}

@@

\TODO{Don't forget to include all the references in this section!}

%FRAM costs more than FRAM -- need to figure out how to execute out of SRAM
New programming and execution models~\cite{dino,ratchet,chain,alpaca} and
architecture mechanisms~\cite{clank, idetic, nvp} help avoid data
inconsistency.  New architectures require costly hardware changes and are
inapplicable to today's systems.  New programming models and compilers force
programmers to use new memory models~\cite{chain,ratchet}, which may inhibit
adoption.  New memory models may also increase the amount of non-volatile
memory accesses by requring non-volatile versioning~\cite{chain} or completely
precluding the use of volatile memory~\cite{ratchet}.  Increasing pressure on
non-volatile memory or precluding the use of volatile memory may degrade
efficiency; e.g., FRAM has a higher access energy and latency than
SRAM~\cite{alpaca,wisp}.  Commercially available devices are often provisioned
with much less SRAM than FRAM~\cite{wolverine,wisp} and effectively leveraging
the relatively high efficiency of SRAM is a key challenge for task-based
intermittent systems.

\TODO{concrete cost multiples here?}

%Task sizing is hard -- how to chop without risking inefficiency or non-term
Several recent programming efforts~\cite{alpaca,chain} proposed {\em
task-based} programming models, that require the programmer to statically
decompose their program into tasks.  A task is like a function with no callers
that can include arbitrary computation and, upon completion, is guaranteed to
have executed {\em atomically} and to have logically {\em committed} memory
updates.  The programmer expresses the flow of control from one task to another
and task control-flow may be input-dependent.  An important programming
challenge in task-based models is that the length of execution of a task is
limited by the total amount of energy that a device can buffer.  Assuming that
input power is negligible compared to operating power, a task will never be
able to complete if its execution consumes more energy than the system can
buffer.  Transitioning from one task to the next imposes a cost and excessively
small tasks may be inefficient.  Appropriately sizing tasks to ensure progress
and high performance is a key challenge for task-based intermittent systems.

In this work, we develop \sys: a new programming and execution model that uses
{\em volatile memory virtualization} to leverage efficient volatile memory and
and {\em dynamic task coalescing} to efficiently execute tasks without
exceeding available energy.  \sys's memory virtualization mechanism uses a
device's SRAM as the working memory during a task's execution, populating the
SRAM with a page of data from FRAM as necessary at the start of the task.  At
the end of the task, the SRAM is committed back to FRAM using efficient DMA
block copies.  \TODO{Brandon sez: Is the next sentence what the system does?
Or are we doing paging?} When a task accesses data outside of a page in SRAM,
the task must fetch data from FRAM.  We explore two strategies. The first
strategy is {\em demand paging}, which swaps the SRAM page with a new page from
FRAM, buffering the swapped-out page until commit.  The second strategy is {\em
buffered direct access} \TODO{better name}, which directly reads and writes
FRAM relying on dynamic double-buffering to ensure memory is consistent at
commit.

\sys's dynamic task coalescing mechanism allows the programmer or a compiler to
intuitively decompose their program into small tasks that amortize fixed
per-task overheads, yet present no risk of exceeding device energy capacity. As
such a decomposition executes, \sys {\em coalesces} dynamically consecutive
tasks. Coalescing elides the commits of coalesced tasks by buffering multiple
tasks' updates in an FRAM commit buffer.  Periodically, as the span of the
coalesced task grows, \sys ends coalescing and commits the state of the
coalesced task.  If a power failure interrupts a sequence of coalesced tasks,
\sys adaptively reduces the number of tasks in that sequence that it will
coalesce, committing sooner in future executions.  Consequently, \sys's
coalescing mechanism allows a program to execute efficiently across a range of
energy buffer sizes, avoiding transition overheads in larger buffers, and
ensuring progress in smaller buffers. \TODO{articulate this last point better}

We evaluated our system on a collection of benchmark programs taken from the
literature, and compare directly to prior work, showing that \sys has high
performance and can flexible target a variety of platforms without recompiling
the code.  We further demonstrate the efficiency and versatility of \sys with
an end-to-end evaluation of a sensing and data processing application deployed
in a lab environment \TODO{blah blah blah eval}.

\TODO{Bullet contribs if there's space}
